### 2022-05-26 paper/localpsf.tex
writing computational cost section of paper.

\paragraph{(1a): Impulse response moments and batches} Computing $V$, $\mu$, and $\Sigma$ require applying $\Aop$ to $1$, $d$, and $d(d+1)/2$ vectors, respectively. Concretely, in one spatial dimensions ($d=1$), computing $V$, $\mu$, and $\Sigma$ requires $1+1+1=3$ operator applies. In two spatial dimensions, this cost is $1+2+3=6$ operator applies. In three spatial dimensions, this cost is $1+3+6=10$ operator applies. Computing each impulse response batch requires applying $\Aop$ to $1$ vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{n_b}$ requires $n_b$ operator applies. In total, computing the impulse response moments and batches requires
\begin{equation*}
	1 + d + d(d+1)/2 + n_b \qquad \text{operator applies.}
\end{equation*}
In a typical application one might have $d=2$ and $n_b=5$, in which case $11$ operator applies would be required. Choosing one sample point batch via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection} may require performing $O(Ns)$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch. Choosing all of the sample points therefore requires performing
\begin{equation*}
	O(Nm) \qquad \text{ellipsoid intersection tests,}
\end{equation*}
where $m$ is the total number of sample points in all batches. The multiplicative dependence of $N$ with $m$ is undesirable, and reducing this cost is a target for future work. The cost of picking sample points could be reduced, for example, by ``thinning'' the sample points. That is, choosing batches from a smaller subset of $N_\text{thin} << N$ points, rather than all $N$ finite element Lagrange nodes. The cost of picking one batch could be reduced from $O(N s)$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked in a batch into an AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection checks. This would reduce the cost of choosing all sample point batches from $O(Nm)$ to $O(N n_b \log s)$. However, in practice the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.

\paragraph{(1b): }

>>>>

\paragraph{(1a): Impulse response moments and batches} Computing $V$, $\mu$, and $\Sigma$ require applying $\Aop$ to $1$, $d$, and $d(d+1)/2$ vectors, respectively. Concretely, in one spatial dimensions ($d=1$), computing $V$, $\mu$, and $\Sigma$ requires $1+1+1=3$ operator applies. In two spatial dimensions, this cost is $1+2+3=6$ operator applies. In three spatial dimensions, this cost is $1+3+6=10$ operator applies. Computing each impulse response batch requires applying $\Aop$ to $1$ vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{n_b}$ requires $n_b$ operator applies. In total, computing the impulse response moments and batches requires
\begin{equation*}
	1 + d + d(d+1)/2 + n_b \qquad \text{operator applies.}
\end{equation*}
In a typical application one might have $d=2$ and $n_b=5$, in which case $11$ operator applies would be required. Choosing one sample point batch via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection} may require performing up to $Ns$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch. Choosing all of the sample points therefore requires performing at most
\begin{equation*}
	Nm \qquad \text{ellipsoid intersection tests,}
\end{equation*}
where $m$ is the total number of sample points in all batches. The multiplicative dependence of $N$ with $m$ is undesirable, and reducing this cost is a target for future work. The cost of picking sample points could be reduced, for example, by ``thinning'' the sample points. That is, choosing batches from a smaller subset of $N_\text{thin} << N$ points, rather than all $N$ finite element Lagrange nodes. The cost of picking one batch could be reduced from $N s$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked in a batch into an AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection checks. This would reduce the cost of choosing all sample point batches from $Nm$ to $O(N n_b \log s)$. However, in practice the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.

\paragraph{(1b): Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(k_h^2 N \log N)$ matrix entries of the approximation, where $k_h$ is the H-matrix rank (maximum rank among the blocks of the H-matrix). To evaluate one matrix entry, first one must perform $k_n$ nearest neighbors queries of the sample points KD tree, and each of these queries requires $O(\log m)$ work. Next, one must finding the mesh cells that $k_n$ distinct points reside in using our mesh AABB tree, which requires $O(k_n \log N)$ work, and evaluate finite element basis functions on each of those cells, which requires $O(k_n)$ work. Finally, one must solve a $k_n \times k_n$ linear system to perform the radial basis function interpolation, which requires $O(k_n^3)$ work. Overall, building the H-matrix therefore requires
\begin{equation*}
	O\left(\left(k_h^2 N \log N\right)\left(k_n \log N + k_n^3\right)\right) \qquad \text{work}.
\end{equation*}
This cost is nearly linear in $N$, scalable with respect to the number of sample points $m$, and quite fast in practice.
